\textbf{Evaluation of machine learning methods for 16S rRNA gene data}

### 2019-03-06 submission/manuscript.Rmd
update

@@

Advances in sequencing technology and decreasing costs of generating 16S rRNA gene sequences have allowed rapid exploration and taxonomic characterization of the human associated microbiome. Currently, the microbiome field is growing at an unprecedented rate and as a result, there is an increasing demand for reproducible methods to identify associations between members of the microbiome and human health. However, this is an undertaking as human associated microbial communities are remarkably uneven. It is unlikely that a single species can explain a disease. Instead, subsets of those communities, in relation to one another and to their host, account for the differences in the health outcomes. Therefore, researchers have started to explore the utility of machine learning (ML) models that use microbiota associated biomarkers to predict human health and to understand the microbial ecology of the human diseases such as liver cirrhosis, colorectal cancer, inflammatory bowel diseases (IBD), obesity, and type 2 diabetes [@zeller_potential_2014; @zackular_human_2014; @baxter_dna_2016; @baxter_microbiota-based_2016; @hale_shifts_2017; @pasolli_machine_2016; @sze_looking_2016; @walters_meta-analyses_2014; @vazquez-baeza_guiding_2018; @qin_alterations_2014; @geman_deep_2018]. However, currently the field's use of ML lacks clarity and consistency on which methods are used and how these methods are implemented. Moreover, there is a lack of deliberation on why a particular ML model is utilized. Recently, there is a trend towards using more complex ML models such as random forest, extreme gradient boosting and neural networks without a discussion on if and how much model interpretibility is necessary for the study [@galkin_human_2018; @reiman_using_2017; @fioravanti_phylogenetic_2017; @ @geman_deep_2018]. If we want to generate reproducible and robust microbiota-associated biomarker models we need to (1) implement consistent, accessible and transparent machine learning practices; (2) use ML models that reflect the goal of the study as it will inform the expectations of model accuracy, complexity, interpretibility, computational efficiency and scalability.

To shed light on how much task definition and modeling choices can affect the outcomes of ML studies, we performed an empirical analysis comparing several different modeling pipelines using the same dataset. We used a previously published colorectal cancer (CRC) study [@baxter_dna_2016] which had fecal 16S rRNA gene sequences and human hemoglobin concentrations from `r paste(nrow(shared))` patients. We built ML models using fecal 16S rRNA gene sequences and human hemoglobin concentrations to predict patients with normal or screen relevant neoplasias (SRN) disease status. The study had `r paste(sum(sum(meta$dx=="normal")+sum(meta$Dx_Bin=="Adenoma")))` normal and `r paste(sum(sum(meta$dx=="cancer")+sum(meta$Dx_Bin=="adv Adenoma")))` SRN samples. We established modeling pipelines for L2-regularized logistic regression, L1 and L2 support vector machines (SVM) with linear and radial basis function kernels, a decision tree, random forest and XGBoost. These models increase in complexity while they decrease in interpretability. Our ML pipeline performed 100 data-splits and utilized held-out test data to evaluate generalization and predicton performance of each ML model. The mean AUROC varied from `r paste(round(test_all[performance_index[1],3], 2))` (std ± `r paste(round(test_all[performance_index[1],4], 2))`) to `r paste(round(test_all[performance_index[7],3], 2))` (std ± `r paste(round(test_all[performance_index[7],4], 2))`). Random Forest and XGBoost had the highest mean AUROC for detecting SRN. Despite the simplicity, the L1-regularized linear kernel SVM performed on par with many of the non-linear models. In terms of computational efficiency, L2 Logistic Regression trained the fastest (`r paste(round(walltime_df[walltime_index[1],2], 3))` hours, std ± `r paste(round(walltime_df[walltime_index[1],3], 3))`), while XGBoost took the longest (`r paste(round(walltime_df[walltime_index[6],2], 3))` hours, std ± `r paste(round(walltime_df[walltime_index[6],3], 3))`). We found that mean cross-validation and testing AUROC could vary by as much as `r paste(round(difference[difference_index[7]], 3))`, which highlights the importance of a seperate held-out test set for evaluation. Aside from evaluating generalization and classification performances for each of these models, this study established standards for modeling pipelines of microbiome-associated machine learning models.

## Results and Discussion

@@

We evaluated the prediction performance of seven binary classification models when applied to held-out test data using AUROC as our metric.  Random Forest and XGBoost had the highest mean AUROC for detecting SRN, `r paste(round(mean(rf$test_aucs), 3))` (std ± `r paste(round(sd(rf$test_aucs), 3))`) and `r paste(round(mean(xgboost$test_aucs), 3))` (std ± `r paste(round(sd(xgboost$test_aucs), 3))`) respectively [Figure 2]. L1 linear SVM and decision tree had significantly lower AUROC values, `r paste(round(mean(l1svm$test_aucs), 3))` (std ± `r paste(round(sd(l1svm$test_aucs), 3))`) and `r paste(round(mean(dt$test_aucs), 3))` (std ± `r paste(round(sd(dt$test_aucs), 3))`)  [Figure 2]. However, they had significanly higher performances than L2 linear SVM, RBF SVM and L2 logistic regression which had mean AUROC values of `r paste(round(mean(l2svm$test_aucs), 3))` (std ± `r paste(round(sd(l2svm$test_aucs), 3))`),  `r paste(round(mean(rbf$test_aucs), 3))` (std ± `r paste(round(sd(rbf$test_aucs), 3))`) and `r paste(round(mean(logit$test_aucs), 3))` (std ± `r paste(round(sd(logit$test_aucs), 3))`) respectively [Figure 2]. We also evaluated the generalization performance of each classifier by comparing their mean cross-validation AUROC and mean testing AUROC. We found a statistically significant difference for classifiers L2 support vector machines (SVM) with linear and radial basis function kernels and L2 logistic regression. These differences were `r paste(round(difference_model[3,1], 3))`, `r paste(round(difference_model[6,1], 3))` and `r paste(round(difference_model[4,1], 3))`, respectively [Figure 2] .

__Hyperparameter optimization for each classifier.__

__The interpretation of classifiers.__

>>>>

# Get the The unpaired two-samples Wilcoxon test to see if models differ from one another signigicantly

rf_xgboost <- wilcox.test(rf$test_aucs, xgboost$test_aucs)
rf_l1svm <- wilcox.test(rf$test_aucs, l1svm$test_aucs)
l1svm_dt <- wilcox.test(l1svm$test_aucs, dt$test_aucs)
l1svm_rbf <- wilcox.test(l1svm$test_aucs, rbf$test_aucs)
l1svm_l2svm <- wilcox.test(l1svm$test_aucs, l2svm$test_aucs)
l1svm_logit <- wilcox.test(l1svm$test_aucs, logit$test_aucs)

# Get the unpaired two-samples Wilcoxon test to see if cv and test differ significantly
logit_diff <- wilcox.test(logit$test_aucs, logit$cv_aucs)
l2svm_diff <- wilcox.test(l2svm$test_aucs, l2svm$cv_aucs)
rbf_diff <- wilcox.test(rbf$test_aucs, rbf$cv_aucs)

@@

Advances in sequencing technology and decreasing costs of generating 16S rRNA gene sequences have allowed rapid exploration of human associated microbiome and its health implications. Currently, the human microbiome field is growing at an unprecedented rate and as a result, there is an increasing demand for methods that identify associations between members of the microbiome and human health. However, this is an undertaking as human associated microbial communities are remarkably complex and uneven. It is unlikely that a single species can explain a disease. Instead, subsets of those communities, in relation to one another and to their host, account for the differences in the health outcomes. Therefore, researchers have started to explore the utility of machine learning (ML) models that use microbiota associated biomarkers to predict human health and to understand the microbial ecology of diseases such as liver cirrhosis, colorectal cancer, inflammatory bowel diseases (IBD), obesity, and type 2 diabetes [@zeller_potential_2014; @zackular_human_2014; @baxter_dna_2016; @baxter_microbiota-based_2016; @hale_shifts_2017; @pasolli_machine_2016; @sze_looking_2016; @walters_meta-analyses_2014; @vazquez-baeza_guiding_2018; @qin_alterations_2014; @geman_deep_2018]. ML methods are effective at recognizing and highlighting patterns in complex microbial datasets. However, currently the field's use of ML lacks clarity and consistency on which methods are used and how these methods are implemented. Moreover, there is a lack of deliberation on why a particular ML model is utilized. Recently, there is a trend towards using more complex ML models such as random forest, extreme gradient boosting and neural networks without a discussion on if and how much model interpretibility is necessary for the study [@galkin_human_2018; @reiman_using_2017; @fioravanti_phylogenetic_2017; @ @geman_deep_2018]. The lack of transparency on methodology and modeling decisions can have negative consequences on model reproducibility and reliability which we need avoid by (1) implementing consistent, accessible and transparent machine learning practices; (2) selecting ML models that reflect the goal of the study as it will inform the expectations of model accuracy, complexity, interpretibility and computational efficiency.

To showcase transparent ML methodologies and to shed light on how much ML model selection can affect modeling results, we performed an empirical analysis comparing several different ML models using the same dataset. We used a previously published colorectal cancer (CRC) study [@baxter_dna_2016] which had fecal 16S rRNA gene sequences and human hemoglobin concentrations from `r paste(nrow(shared))` patients. We built ML models using fecal 16S rRNA gene sequences and human hemoglobin concentrations to predict patients with normal or screen relevant neoplasias (SRN) disease status. The study had `r paste(sum(sum(meta$dx=="normal")+sum(meta$Dx_Bin=="Adenoma")))` normal and `r paste(sum(sum(meta$dx=="cancer")+sum(meta$Dx_Bin=="adv Adenoma")))` SRN samples. We established modeling pipelines for L2-regularized logistic regression, L1 and L2 support vector machines (SVM) with linear and radial basis function kernels, a decision tree, random forest and XGBoost which increase in complexity and decrease in interpretability respectively. Our ML pipeline performed 100 data-splits and utilized held-out test data to evaluate generalization and predicton performance of each ML model. The mean AUROC varied from `r paste(round(test_all[performance_index[1],3], 2))` (std ± `r paste(round(test_all[performance_index[1],4], 2))`) to `r paste(round(test_all[performance_index[7],3], 2))` (std ± `r paste(round(test_all[performance_index[7],4], 2))`). Random Forest and XGBoost had the highest mean AUROC for detecting SRN. Despite the simplicity, the L1-regularized linear kernel SVM followed Random Forest and XGBoost in performance. In terms of computational efficiency, L2 SVM with linear kernel trained the fastest (`r paste(round(walltime_df[walltime_index[1],2], 3))` hours, std ± `r paste(round(walltime_df[walltime_index[1],3], 3))`), while XGBoost took the longest (`r paste(round(walltime_df[walltime_index[6],2], 3))` hours, std ± `r paste(round(walltime_df[walltime_index[6],3], 3))`). We found that mean cross-validation and testing AUROC could vary by as much as `r paste(round(difference[difference_index[7]], 3))`, which highlights the importance of a seperate held-out test set for evaluation. Aside from evaluating generalization and classification performances for each of these models, this study established standards for modeling pipelines of microbiome-associated machine learning models.

## Results

__Model selection and construction__

We established modeling pipelines for binary classification with L2-regularized logistic regression, L1 and L2 support vector machines (SVM) with linear and radial basis function kernels, a decision tree, random forest and XGBoost to emphasize the differences in  model accuracy, complexity, interpretibility, computational efficiency and scalability due to  model selection. We randomly split the data into stratified training and test sets 100 times with a 80/20 split [Figure 1]. The training set consisted of 393 patients (209 SRN), while the test set was composed of 97 patients (52 SRN). The training data was used for training purposes and validation of parameter selection, and the test set was used for evaluation purposes. Validation of parameter selection was repeated 100 times to get a robust parameter selection for each model [Figure 1].

@@

We evaluated the prediction performance of seven binary classification models when applied to held-out test data over 100 data-splits using AUROC as classification performance metric. Random Forest and XGBoost had the highest mean AUROC for detecting SRN, `r paste(round(mean(rf$test_aucs), 3))` (std ± `r paste(round(sd(rf$test_aucs), 3))`) and `r paste(round(mean(xgboost$test_aucs), 3))` (std ± `r paste(round(sd(xgboost$test_aucs), 3))`) respectively [Figure 2]. L1 linear SVM and decision tree had significantly lower AUROC values, `r paste(round(mean(l1svm$test_aucs), 3))` (std ± `r paste(round(sd(l1svm$test_aucs), 3))`) and `r paste(round(mean(dt$test_aucs), 3))` (std ± `r paste(round(sd(dt$test_aucs), 3))`)  [Figure 2]. However, they had significanly higher performances than L2 linear SVM, RBF SVM and L2 logistic regression which had mean AUROC values of `r paste(round(mean(l2svm$test_aucs), 3))` (std ± `r paste(round(sd(l2svm$test_aucs), 3))`),  `r paste(round(mean(rbf$test_aucs), 3))` (std ± `r paste(round(sd(rbf$test_aucs), 3))`) and `r paste(round(mean(logit$test_aucs), 3))` (std ± `r paste(round(sd(logit$test_aucs), 3))`) respectively [Figure 2]. We also evaluated the generalization performance of each classifier by comparing their mean cross-validation AUROC and mean testing AUROC. We found a statistically significant difference for classifiers L2 support vector machines (SVM) with linear and radial basis function kernels and L2 logistic regression (p-values = `r paste(format(l2svm_diff$p.value, digits=2))`, `r paste(format(rbf_diff$p.value, digits=2))` and `r paste(format(logit_diff$p.value, digits=2))` respectively). These differences were `r paste(round(difference_model[3,1], 3))`, `r paste(round(difference_model[6,1], 3))` and `r paste(round(difference_model[4,1], 3))`, respectively [Figure 2] .

__The complexity and interpretibility of each classifier.__

We interpreted the feature weights of L1 and L2 SVM with linear kernel and log odds ratios of L2 logistic regression using the training data. To get a sense of the importance of each feature we randomly subsampled 80% of the data 100 times and trained 100 models; this resulted in 100 different sets of feature weights. We reported the mean weights of the 10 most important features for[Figure 4]. FIT levels were

__The computational efficiency of each classifier.__

Linear models trained faster than non-linear models. L2 logistic Rregression and L1 and L2 SVM with linear kernel had training times of `r paste(60*(round(walltime_df[walltime_index[1],2], 2)))` min, (std ± `r paste(60*(round(walltime_df[walltime_index[1],3], 2)))`), `r paste(60*(round(walltime_df[walltime_index[2],2], 2)))` min, (std ± `r paste(60*(round(walltime_df[walltime_index[2],3], 2)))`) and `r paste(60*(round(walltime_df[walltime_index[3],2], 2)))` min, (std ± `r paste(60*(round(walltime_df[walltime_index[3],3], 2)))`) respectively. Whereas, SVM with radial basis function kernel, decision tree, random forest and xgboost had training times of `r paste(round(walltime_df[walltime_index[4],2], 2))` hrs, (std ± `r paste(round(walltime_df[walltime_index[4],3], 2))`), `r paste((round(walltime_df[walltime_index[5],2], 2)))` hrs, (std ± `r paste((round(walltime_df[walltime_index[5],3], 2)))`), `r paste(round(walltime_df[walltime_index[6],2], 2))` hrs, (std ± `r paste(round(walltime_df[walltime_index[6],3], 2))`) and `r paste(round(walltime_df[walltime_index[7],2], 2))` hrs, (std ± `r paste(round(walltime_df[walltime_index[7],3], 2))`), respectively.

### 2019-03-06 submission/manuscript.tex
update

Advances in sequencing technology and decreasing costs of generating 16S
rRNA gene sequences have allowed rapid exploration and taxonomic
characterization of the human associated microbiome. Currently, the
microbiome field is growing at an unprecedented rate and as a result,
there is an increasing demand for reproducible methods to identify
associations between members of the microbiome and human health.
However, this is an undertaking as human associated microbial
communities are remarkably uneven. It is unlikely that a single species
can explain a disease. Instead, subsets of those communities, in
relation to one another and to their host, account for the differences
in the health outcomes. Therefore, researchers have started to explore
the utility of machine learning (ML) models that use microbiota
associated biomarkers to predict human health and to understand the
microbial ecology of the human diseases such as liver cirrhosis,
colorectal cancer, inflammatory bowel diseases (IBD), obesity, and type
2 diabetes (1--11). However, currently the field's use of ML lacks
clarity and consistency on which methods are used and how these methods
are implemented. Moreover, there is a lack of deliberation on why a
particular ML model is utilized. Recently, there is a trend towards
using more complex ML models such as random forest, extreme gradient
boosting and neural networks (11--14) without a discussion on if and how
much model interpretibility is necessary for the study. If we want to
generate reproducible and robust microbiota-associated biomarker models
we need to (1) implement consistent, accessible and transparent machine
learning practices; (2) use ML models that reflect the goal of the study
as it will inform the expectations of model accuracy, complexity,
interpretibility, computational efficiency and scalability.

To shed light on how much task definition and modeling choices can
affect the outcomes of ML studies, we performed an empirical analysis
comparing several different modeling pipelines using the same dataset.
We used a previously published colorectal cancer (CRC) study (3) which
had fecal 16S rRNA gene sequences and human hemoglobin concentrations
from 490 patients. We built ML models using fecal 16S rRNA gene
sequences and human hemoglobin concentrations to predict patients with
normal or screen relevant neoplasias (SRN) disease status. The study had
261 normal and 229 SRN samples. We established modeling pipelines for
L2-regularized logistic regression, L1 and L2 support vector machines
(SVM) with linear and radial basis function kernels, a decision tree,
random forest and XGBoost. These models increase in complexity while
they decrease in interpretability. Our ML pipeline performed 100
data-splits and utilized held-out test data to evaluate generalization
and predicton performance of each ML model. The mean AUROC varied from
0.67 (std ± 0.05) to 0.82 (std ± 0.04). Random Forest and XGBoost had
the highest mean AUROC for detecting SRN. Despite the simplicity, the
L1-regularized linear kernel SVM performed on par with many of the
non-linear models. In terms of computational efficiency, L2 Logistic
Regression trained the fastest (0.105 hours, std ± 0.013), while XGBoost
took the longest (82.47 hours, std ± 17.045). We found that mean
cross-validation and testing AUROC could vary by as much as 0.062, which
highlights the importance of a seperate held-out test set for
evaluation. Aside from evaluating generalization and classification
performances for each of these models, this study established standards
for modeling pipelines of microbiome-associated machine learning models.

\subsection{Results and Discussion}\label{results-and-discussion}

@@

We evaluated the prediction performance of seven binary classification
models when applied to held-out test data using AUROC as our metric.
Random Forest and XGBoost had the highest mean AUROC for detecting SRN,
0.816 (std ± 0.039) and 0.814 (std ± 0.038) respectively {[}Figure 2{]}.
L1 linear SVM and decision tree had significantly lower AUROC values,
0.748 (std ± 0.043) and 0.741 (std ± 0.038) {[}Figure 2{]}. However,
they had significanly higher performances than L2 linear SVM, RBF SVM
and L2 logistic regression which had mean AUROC values of 0.67 (std ±
0.054), 0.684 (std ± 0.049) and 0.677 (std ± 0.053) respectively
{[}Figure 2{]}. We also evaluated the generalization performance of each
classifier by comparing their mean cross-validation AUROC and mean
testing AUROC. We found a statistically significant difference for
classifiers L2 support vector machines (SVM) with linear and radial
basis function kernels and L2 logistic regression. These differences
were 0.062, 0.046 and 0.061, respectively {[}Figure 2{]} .

\textbf{Hyperparameter optimization for each classifier.}

\textbf{The interpretation of classifiers.}

>>>>

Advances in sequencing technology and decreasing costs of generating 16S
rRNA gene sequences have allowed rapid exploration of human associated
microbiome and its health implications. Currently, the human microbiome
field is growing at an unprecedented rate and as a result, there is an
increasing demand for methods that identify associations between members
of the microbiome and human health. However, this is an undertaking as
human associated microbial communities are remarkably complex and
uneven. It is unlikely that a single species can explain a disease.
Instead, subsets of those communities, in relation to one another and to
their host, account for the differences in the health outcomes.
Therefore, researchers have started to explore the utility of machine
learning (ML) models that use microbiota associated biomarkers to
predict human health and to understand the microbial ecology of diseases
such as liver cirrhosis, colorectal cancer, inflammatory bowel diseases
(IBD), obesity, and type 2 diabetes (1--11). ML methods are effective at
recognizing and highlighting patterns in complex microbial datasets.
However, currently the field's use of ML lacks clarity and consistency
on which methods are used and how these methods are implemented.
Moreover, there is a lack of deliberation on why a particular ML model
is utilized. Recently, there is a trend towards using more complex ML
models such as random forest, extreme gradient boosting and neural
networks without a discussion on if and how much model interpretibility
is necessary for the study (11--14). The lack of transparency on
methodology and modeling decisions can have negative consequences on
model reproducibility and reliability which we need avoid by (1)
implementing consistent, accessible and transparent machine learning
practices; (2) selecting ML models that reflect the goal of the study as
it will inform the expectations of model accuracy, complexity,
interpretibility and computational efficiency.

To showcase transparent ML methodologies and to shed light on how much
ML model selection can affect modeling results, we performed an
empirical analysis comparing several different ML models using the same
dataset. We used a previously published colorectal cancer (CRC) study
(3) which had fecal 16S rRNA gene sequences and human hemoglobin
concentrations from 490 patients. We built ML models using fecal 16S
rRNA gene sequences and human hemoglobin concentrations to predict
patients with normal or screen relevant neoplasias (SRN) disease status.
The study had 261 normal and 229 SRN samples. We established modeling
pipelines for L2-regularized logistic regression, L1 and L2 support
vector machines (SVM) with linear and radial basis function kernels, a
decision tree, random forest and XGBoost which increase in complexity
and decrease in interpretability respectively. Our ML pipeline performed
100 data-splits and utilized held-out test data to evaluate
generalization and predicton performance of each ML model. The mean
AUROC varied from 0.68 (std ± 0.05) to 0.82 (std ± 0.04). Random Forest
and XGBoost had the highest mean AUROC for detecting SRN. Despite the
simplicity, the L1-regularized linear kernel SVM followed Random Forest
and XGBoost in performance. In terms of computational efficiency, L2 SVM
with linear kernel trained the fastest (0.127 hours, std ± 0.01), while
XGBoost took the longest (82.47 hours, std ± 17.045). We found that mean
cross-validation and testing AUROC could vary by as much as 0.061, which
highlights the importance of a seperate held-out test set for
evaluation. Aside from evaluating generalization and classification
performances for each of these models, this study established standards
for modeling pipelines of microbiome-associated machine learning models.

\subsection{Results}\label{results}

\textbf{Model selection and construction}

We established modeling pipelines for binary classification with
L2-regularized logistic regression, L1 and L2 support vector machines
(SVM) with linear and radial basis function kernels, a decision tree,
random forest and XGBoost to emphasize the differences in model
accuracy, complexity, interpretibility, computational efficiency and
scalability due to model selection. We randomly split the data into
stratified training and test sets 100 times with a 80/20 split {[}Figure
1{]}. The training set consisted of 393 patients (209 SRN), while the
test set was composed of 97 patients (52 SRN). The training data was
used for training purposes and validation of parameter selection, and
the test set was used for evaluation purposes. Validation of parameter
selection was repeated 100 times to get a robust parameter selection for
each model {[}Figure 1{]}.

@@

We evaluated the prediction performance of seven binary classification
models when applied to held-out test data over 100 data-splits using
AUROC as classification performance metric. Random Forest and XGBoost
had the highest mean AUROC for detecting SRN, 0.816 (std ± 0.039) and
0.814 (std ± 0.04) respectively {[}Figure 2{]}. L1 linear SVM and
decision tree had significantly lower AUROC values, 0.75 (std ± 0.045)
and 0.741 (std ± 0.038) {[}Figure 2{]}. However, they had significanly
higher performances than L2 linear SVM, RBF SVM and L2 logistic
regression which had mean AUROC values of 0.68 (std ± 0.052), 0.684 (std
± 0.049) and 0.677 (std ± 0.053) respectively {[}Figure 2{]}. We also
evaluated the generalization performance of each classifier by comparing
their mean cross-validation AUROC and mean testing AUROC. We found a
statistically significant difference for classifiers L2 support vector
machines (SVM) with linear and radial basis function kernels and L2
logistic regression (p-values = 2.4e-16, 3.5e-13 and 9.4e-20
respectively). These differences were 0.052, 0.047 and 0.061,
respectively {[}Figure 2{]} .

\textbf{The complexity and interpretibility of each classifier.}

\textbf{The computational efficiency of each classifier.}

Linear models trained faster than non-linear models. L2 logistic
Rregression and L1 and L2 SVM with linear kernel had training times of
7.8 min, (std ± 0.6), 9.6 min, (std ± 0.6) and 13.2 min, (std ± 1.2)
respectively. Whereas, SVM with radial basis function kernel, decision
tree, random forest and xgboost had training times of 2.21 hrs, (std ±
0.25), 4.2 hrs, (std ± 0.49), 82.47 hrs, (std ± 17.05) and 116.83 hrs,
(std ± 17.24), respectively.
